Dark Control:\\ Towards a Unified Account of Default Mode Function\\ by Markov Decision Processes

### 2017-12-04 paper/paper.tex
typos pointed-out by rev #2

RL is an area of machine learning concerned with learning optimal behavioral through interactions
with an \textit{environment}, the goal being to maximize some notion of \textit{cumulative reward}.
The optimal behavior typical
takes the future into account, as some rewards could be \textit{delayed}.
Through repeated interactions with the environment,
the agent learns to reach goals and optimize reward signals
in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
At a given moment, each taken \textit{action} $a$ triggers a change
in the \textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
$r = r(s, a,s')$ collected by the agent.
If the collected reward outcome yields a negative value it can be
more naturally interpreted as \textit{punishment}.
In this view, the environment is partially controlled by
the action of the agent and the reward can be thought
of as satisfaction \textemdash or aversion \textemdash accompanying the execution of
a particular action.
% Action may be delayed to achieve substantial increases in expected reward
% that is expected to grow with time.

@@

At the current step $k$, computing an optimal parameter update then corresponds to
finding the model parameters $\theta_{k}$ which minimize the following mean-squared loss function
\begin{equation}
  \mathcal L(\theta^Q_{k})
  % = \mathbb E_{(s, a, r, s') \sim \mathcal D}[(\tilde{Q}(s, a|\theta) - \tilde{Q}(s',a'|\theta^Q_{k}))^2]
  = \mathbb E_{(s, a, r, s') \sim \mathcal D}\left[\frac{1}{2}(\tilde{Q}(s, a|\theta_{k}) - y_k)^2\right],
  \label{eq:oracle}
\end{equation}
where $y_k$ is defined in \eqref{eq:bellman}.
% \textbf{We still may need to explain y in the formula with one sentence.}
A recently proposed, practically successful alternative approach is to learn this
representation using an artificial deep neural-network model, leading to the
so-called \textit{deep Q-learning}~\citep{mnih2015,silver2016mastering} family of methods which
is the current state-of-the-art in RL research.
The set of model parameters $\theta$ that instantiate the non-linear interactions
between layers of the artificial neural network
may find a neurobiological correspondence in the adaptive strengths of axonal
connections between neurons from the different levels
of the neural processing hierarchy
\citep{mesulam1998sensation, taylor2015global}.

@@

\begin{mdframed}
  \vspace{1em}
  \textbf{Box 2: Hypotheses for testing the MDP account of DMN function}
\begin{enumerate}
\item Experiment (Humans): We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry. During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigms. We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, and the more the fMRI signals in the reward circuits should be independent of the reward context
in the sensory environment.

@@

In particular,
environmental stimuli especially important for humans are frequently of
social nature. This is not surprising
given that
the complexity of the social systems
is likely to be a human-defining property
\citep{tomasello2009cultural}.
According to the ``social brain hypothesis'',
the human brain has especially been shaped for
forming and maintaining increasingly complex
social systems,
which allows solving ecological problems by means of social relationships
\citep{whiten1988machiavellian}.
Indeed, social topics amounted to roughly
two thirds of human everyday communication \citep{dunbar1997human},
while
mind-wandering at daytime and dreams during sleep
are rich in stories about people and
the complex relationships between them.
%
In line with this, the DMN was argued to be specialized in
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default},
cats \citep{popa2009contrasting},
and rats \citep{lu2012rat}, three species with
social-cognitive capacities that are supposedly less advanced than in humans.

@@

It is an important feature of the proposed artificial intelligence perspective
on DMN biology that it is practically computable and
yields falsifiable neuroscientific hypotheses.
%
Neuroscience experiments could be designed that operationalize
the set of action, value, and state variables that govern
the behavior of intelligent RL agents.
%
At the least, we propose an alternative vocabulary to
describe, contextualize, and interpret experimental findings in neuroscience studies
on the DMN.
%
Ultimately,
DMN activity may instantiate a holistic integration
ranging from real experience over purposeful dreams to anticipated futures
for continued refinement of the organism's fate. Indeed, a major hurdle in explaining DMN activity from brain-imaging studies has been its very similar engaged across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., dreams during sleep or daytime mind-wandering), and thinking about future scenarios (e.g., prospection or delay discounting). The MDP account of DMN function offers a natural integration of these a-priori diverging mental processes into a common framework. % It is this feature of our functional view that we underline in the referred sentence.

>>>>

RL is an area of machine learning concerned with learning optimal behavioral through interactions
with an \textit{environment}, the goal being to maximize some notion of \textit{cumulative reward}.
The optimal behavior typically
takes the future into account, as some rewards could be \textit{delayed}.
Through repeated interactions with the environment,
the agent learns to reach goals and optimize reward signals
in an iterative trial-and-error fashion (Fig. \ref{fig:rl}).
At a given moment, each taken \textit{action} $a$ triggers a change
in the \textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
$r = r(s, a,s')$ collected by the agent.
If the collected reward outcome yields a negative value it can be
more naturally interpreted as \textit{punishment}.
In this view, the environment is partially controlled by
the action of the agent and the reward can be thought
of as satisfaction \textemdash or aversion \textemdash accompanying the execution of
a particular action.
% Action may be delayed to achieve substantial increases in expected reward
% that is expected to grow with time.

@@

At the current step $k$, computing an optimal parameter update then corresponds to
finding the model parameters $\theta_{k}$ which minimize the following mean-squared loss function
\begin{equation}
  \mathcal L(\theta^Q_{k})
  % = \mathbb E_{(s, a, r, s') \sim \mathcal D}[(\tilde{Q}(s, a|\theta) - \tilde{Q}(s',a'|\theta^Q_{k}))^2]
  = \mathbb E_{(s, a, r, s') \sim \mathcal D}\left[\frac{1}{2}(\tilde{Q}(s, a|\theta_{k}) - y_k)^2\right],
  \label{eq:oracle}
\end{equation}
where $y_k$ is defined in \eqref{eq:bellman}.
% \textbf{We still may need to explain y in the formula with one sentence.}
A recently proposed, practically successful alternative approach is, to learn this
representation using an artificial deep neural-network model, leading to the
so-called \textit{deep Q-learning}~\citep{mnih2015,silver2016mastering} family of methods which
is the current state-of-the-art in RL research.
The set of model parameters $\theta$ that instantiate the non-linear interactions
between layers of the artificial neural network
may find a neurobiological correspondence in the adaptive strengths of axonal
connections between neurons from the different levels
of the neural processing hierarchy
\citep{mesulam1998sensation, taylor2015global}.

@@

\begin{mdframed}
  \vspace{1em}
  \textbf{Box 2: Hypotheses for testing the MDP account of DMN function}
\begin{enumerate}
\item Experiment (Humans): We hypothesize a functional relationship between the DMN closely associated with the occurrence of stimulus-independent thoughts and the reward circuitry. During an iterative neuroeconomic two-player game, fMRI signals in the DMN could be used to predict reward-related signals in the nucleus accumbens across trials in a continuous learning paradigm. We expect that the more DMN activity is measured to be increased, supposedly the higher the tendency for stimulus-independent thoughts, and the more the fMRI signals in the reward circuits should be independent of the reward context
in the sensory environment.

@@

In particular,
environmental stimuli are especially important for humans are frequently of
social nature. This is not surprising
given that
the complexity of the social systems
is likely to be a human-defining property
\citep{tomasello2009cultural}.
According to the ``social brain hypothesis'',
the human brain has especially been shaped for
forming and maintaining increasingly complex
social systems,
which allows solving ecological problems by means of social relationships
\citep{whiten1988machiavellian}.
Indeed, social topics amounted to roughly
two thirds of human everyday communication \citep{dunbar1997human},
while
mind-wandering at daytime and dreams during sleep
are rich in stories about people and
the complex relationships between them.
%
In line with this, the DMN was argued to be specialized in
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default},
cats \citep{popa2009contrasting},
and rats \citep{lu2012rat}, three species with
social-cognitive capacities that are supposedly less advanced than in humans.

@@

It is an important feature of the proposed artificial intelligence perspective
on DMN biology that it is practically computable and
yields falsifiable neuroscientific hypotheses.
%
Neuroscience experiments could be designed that operationalize
the set of action, value, and state variables that govern
the behavior of intelligent RL agents.
%
At the least, we propose an alternative vocabulary to
describe, contextualize, and interpret experimental findings in neuroscience studies
on the DMN.
%
Ultimately,
DMN activity may instantiate a holistic integration
ranging from real experience over purposeful dreams to anticipated futures
for continued refinement of the organism's fate. Indeed, a major hurdle in explaining DMN activity from brain-imaging studies has been its very similar engagement across time scales: thinking about the past (e.g., autobiographical memory), thinking about hypothetical presents (e.g., dreams during sleep or daytime mind-wandering), and thinking about future scenarios (e.g., prospection or delay discounting). The MDP account of DMN function offers a natural integration of these a-priori diverging mental processes into a common framework. % It is this feature of our functional view that we underline in the referred sentence.
